---
title: "Tomato nutrient study correlations, 2nd version"
author: "Rachel Schattman and Becky Maden"
date: "November 27, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library
```{r}
library(dplyr)
library(ggpubr)
library(ggplot2)
library(car)
library(gvlma)
library(reshape2)
```
## Questions of interest

1.	The effects of cumulative nutrient applications on leaf tissue and soil nutrient levels the month after. (MASS only)
2.	The effects of soil test in any given month and the leaf tissue samples of the following month.
3.	The effects of diseases on yield (H1: powdery mildew will have the biggest effect)
4.	The effect of compaction on yield
5.	The effect of plant density on yield
6.	The effect of planting date on yield


## Start script
# Data preparation

```{r}
script_path <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV"
in_dir <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV"
out_dir <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/output/CSV"
```

# read in datasets
```{r}
FARMS <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/Farms.csv"))
NUTADDED <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/NutAdd.csv"))
NUTLEVELS <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/NutLevel.csv"))
PESTS <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/Pests.csv"))
SPACEYIELD <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/SpaceYield.csv"))
```

## Merge dataframes
```{r}
NUTS <-merge(NUTADDED, SPACEYIELD)
NUTS2 <- merge(NUTS, NUTLEVELS)
```


# Question 2: The effects of soil test in any given month and the leaf tissue samples of the following month.

## subset modified morgans (MM), saturdated media (SM) and leaf tissue (LT) tests by month after transplan (MX in SampPer). MM and SM are for same month, leaf tissue is for following month.
```{r}
# subset 1M rows, MM and SM columns
MM_SM_M1 <- NUTS2[which(NUTS2$SampPer == "1M"), names(NUTS2) %in% c("ID", "TMY", "MYsqFT", "NSME", "ASME", "PSME", "KSME", "MagSME", "CaSME", "AlSME", "BoSME", "CoSME", "FeSME", "MangSME", "SoSME", "SuSME", "ZSME", "NMM", "PMM", "KMM", "CaMM", "MaMM",  "SuMM", "BoMM", "CoMM", "FeMM", "MangMM", "ZMM")] 

# subset 2M rows, LT columns
LA_M2 <- NUTS2[which(NUTS2$SampPer == "2M"), names(NUTS2) %in% c("TNLA", "CALA", "KLA", "MgLA", "PLA", "AlLA", "BLA", "CuLA", "FeLA", "MnLA",  "ZLA")]

# merge MM_SM_M1 and LA_M2
M1_M2 <- merge(MM_SM_M1, LA_M2)

# some columns need to be tranformed into numberic vectors ### well this didn't work. 
as.numeric(M1_M2$MangSME)
is.numeric(M1_M2$MangSME)

```

## test variables for normal distribution
```{r}
shaptest<-lapply(M1_M2, shapiro.test)
shapiro.test(M1_M2$TMY)
shapiro.test(M1_M2$MYsqFT)
shapiro.test(M1_M2$NSME)
shapiro.test(M1_M2$ASME)
shapiro.test(M1_M2$PSME)
shapiro.test(M1_M2$KSME)
shapiro.test(M1_M2$MagSME) #not normal dist
shapiro.test(M1_M2$CaSME)
shapiro.test(M1_M2$AlSME)
shapiro.test(M1_M2$BoSME)
shapiro.test(M1_M2$CoSME)
shapiro.test(M1_M2$FeSME) #not normal dist


shapiro.test(M1_M2$MangSME) 
shapiro.test(M1_M2$SoSME)
shapiro.test(M1_M2$SuSME)
shapiro.test(M1_M2$ZSME)
shapiro.test(M1_M2$NMM)
shapiro.test(M1_M2$PMM)
shapiro.test(M1_M2$KMM)
shapiro.test(M1_M2$CaMM)
shapiro.test(M1_M2$MaMM)
shapiro.test(M1_M2$SuMM)
shapiro.test(M1_M2$BoMM)
shapiro.test(M1_M2$CoMM)
shapiro.test(M1_M2$FeMM)
shapiro.test(M1_M2$MangMM)
shapiro.test(M1_M2$ZMM)
shapiro.test(M1_M2$TNLA)
shapiro.test(M1_M2$CALA)
shapiro.test(M1_M2$KLA)
shapiro.test(M1_M2$MgLA)
shapiro.test(M1_M2$PLA)
shapiro.test(M1_M2$AlLA)
shapiro.test(M1_M2$BLA)
shapiro.test(M1_M2$CuLA)
shapiro.test(M1_M2$FeLA)
shapiro.test(M1_M2$MnLA)
shapiro.test(M1_M2$ZLA)
```

## Regress leaf tissue analysis on SME and MM for each nutrient
```{r}

```









# Nitrate
## Create dataframe w/variables of interest for Nitrate only, omit rows with NA


```{r}

tomatofun_nitrate <- subset(data, select = c("Nitrate.N..ppm..SME", "Nitrate.N..ppm..MM", "TN.....Leaf.Analysis"))
tomatofun_nitrateNONA <- na.omit(tomatofun_nitrate) 

```

## Check distribution - all are non normal

```{r}
shapiro.test(tomatofun_nitrateNONA$Nitrate.N..ppm..SME)
shapiro.test(tomatofun_nitrateNONA$Nitrate.N..ppm..MM)
shapiro.test(tomatofun_nitrateNONA$TN.....Leaf.Analysis)
```
## Correlation tests

```{r}

cor(tomatofun_nitrateNONA, method = "spearman")

cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..SME, y = tomatofun_nitrateNONA$Nitrate.N..ppm..MM, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..SME, y = tomatofun_nitrateNONA$TN.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..MM, y = tomatofun_nitrateNONA$TN.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)

```
## plot SME and MM correlations
```{r}
?plot
plot(tomatofun_nitrateNONA$Nitrate.N..ppm..SME, tomatofun_nitrateNONA$Nitrate.N..ppm..MM, type = "p", main = "Nitrate SME and MM, p < 0.001", xlab = "SME PPM", ylab = "MM PPM")
abline(lm(tomatofun_nitrateNONA$Nitrate.N..ppm..SME~tomatofun_nitrateNONA$Nitrate.N..ppm..MM), col = "red") # regression line
```
# Phosphorus
## Create dataframe w/variables of interest for P only, omit rows with NA

```{r}
tomatofun_P <- subset(data, select = c("Phosphorus..ppm..SME", "Phosphorus..lb.Ac..MM", "P.....Leaf.Analysis"))
tomatofun_P_NONA <- na.omit(tomatofun_P) 

```

## Check distribution 

```{r}
shapiro.test(tomatofun_P_NONA$Phosphorus..ppm..SME)
shapiro.test(tomatofun_P_NONA$Phosphorus..lb.Ac..MM)
shapiro.test(tomatofun_P_NONA$P.....Leaf.Analysis)
```

## Correlation tests

```{r}

cor(tomatofun_P_NONA, method = "spearman")

cor.test(x = tomatofun_P_NONA$Phosphorus..ppm..SME, y = tomatofun_P_NONA$Phosphorus..lb.Ac..MM, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_P_NONA$Phosphorus..ppm..SME, y = tomatofun_P_NONA$P.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_P_NONA$Phosphorus..lb.Ac..MM, y = tomatofun_P_NONA$P.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)

```

## plot SME and MM correlations
```{r}

plot(tomatofun_P_NONA$Phosphorus..lb.Ac..MM, tomatofun_P_NONA$Phosphorus..ppm..SME, type = "p", main = "Phosphorus SME and MM, p < 0.001", xlab = "MM PPM", ylab = "SME PPM")

abline(lm(tomatofun_P_NONA$Phosphorus..ppm..SME~tomatofun_P_NONA$Phosphorus..lb.Ac..MM), col = "red") # regression line
```

# Potassium
## Create dataframe w/variables of interest for P only, omit rows with NA
head(data)

```{r}
tomatofun_K <- subset(data, select = c("Potassium..ppm..SME", "Potassium..lb.Ac..MM", "K.....Leaf.Analysis"))
tomatofun_K_NONA <- na.omit(tomatofun_K) 

```

## Check distribution 

```{r}
shapiro.test(tomatofun_K_NONA$Potassium..ppm..SME)
shapiro.test(tomatofun_K_NONA$Potassium..lb.Ac..MM)
shapiro.test(tomatofun_K_NONA$K.....Leaf.Analysis)
```

```{r}

cor(tomatofun_K_NONA, method = "spearman")

cor.test(x = tomatofun_K_NONA$Potassium..ppm..SME, y = tomatofun_K_NONA$Potassium..lb.Ac..MM, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_K_NONA$Potassium..ppm..SME, y = tomatofun_K_NONA$K.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_K_NONA$Potassium..lb.Ac..MM, y = tomatofun_K_NONA$K.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)

```

## plot SME and MM correlations
```{r}

plot(tomatofun_K_NONA$Potassium..ppm..SME, tomatofun_K_NONA$Potassium..lb.Ac..MM, type = "p", main = "Potassium SME and MM, p < 0.001", xlab = "SME PPM", ylab = "MM PPM")
abline(lm(tomatofun_K_NONA$Potassium..lb.Ac..MM~tomatofun_K_NONA$Potassium..ppm..SME), col = "red") # regression line
```

# P SME and pH MM
## Create dataframe w/variables of interest, omit rows with NA
head(data)
```{r}
tomatofun_P_pH <- subset(data, select = c("Phosphorus..ppm..SME", "Soil.pH"))
tomatofun_P_pH_NONA <- na.omit(tomatofun_P_pH) 
```

```{r}

cor(tomatofun_P_pH_NONA, method = "spearman")

cor.test(x = tomatofun_P_pH_NONA$Phosphorus..ppm..SME, y = tomatofun_P_pH_NONA$Soil.pH, alternative = "two.sided", method = "spearman", exact = NULL)

plot(tomatofun_P_pH_NONA$Phosphorus..ppm..SME, tomatofun_P_pH_NONA$Soil.pH, type = "p", main = "Phosphorus SME and pH, p < 0.001", xlab = "SME PPM", ylab = "pH")
abline(lm(tomatofun_P_pH_NONA$Soil.pH~tomatofun_P_pH_NONA$Phosphorus..ppm..SME), col = "red") # regression line
```

# Take home message: No significant difference between Nitrate levels (any test) by sample period. 
```{r}

tomatofun_yield_N <- subset(data, select = c("Total.Yield.per.sq.ft", "Sample.Period", "Nitrate.N..ppm..SME", "Nitrate.N..ppm..MM", "TN.....Leaf.Analysis"))

tomatofun_yield_N_NONA <- na.omit(tomatofun_yield_N)

#set comparisons for error plots
my_comparisons <- list(c("Pre-Plant", "1-MAT"), c("Pre-Plant", "2-MAT"), c("Pre-Plant", "3-MAT"), c("Pre-Plant", "4-MAT"), c("Pre-Plant", "5-MAT"))
my_comparisonsb <- list(c("1-MAT", "2-MAT"), c("1-MAT", "3-MAT"), c("1-MAT", "4-MAT"), c("1-MAT", "5-MAT"))

#correct order of sample periods
tomatofun_yield_N_NONA$"Sample.Period" <- factor(tomatofun_yield_N_NONA$"Sample.Period", levels = c('Pre-Plant', 
                                                                                           '1-MAT', 
                                                                                            '2-MAT', 
                                                                                            '3-MAT', 
                                                                                            '4-MAT', 
                                                                                            '5-MAT'))

ggerrorplot(tomatofun_yield_N_NONA, x = "Sample.Period", 
            y = "Nitrate.N..ppm..SME",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and SME nitrate",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-500, 1000),
            legend = "bottom",
            xlab = "sample period",
            ylab = "SME Nitrate",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -250, label.x = )

ggerrorplot(tomatofun_yield_N_NONA, x = "Sample.Period", 
            y = "Nitrate.N..ppm..MM",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and MM nitrate",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-500, 1000),
            legend = "bottom",
            xlab = "sample period",
            ylab = "MM Nitrate",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -250, label.x = )

ggerrorplot(tomatofun_yield_N_NONA, x = "Sample.Period", 
            y = "TN.....Leaf.Analysis",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and total N(leaf)",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-2, 8),
            legend = "bottom",
            xlab = "sample period",
            ylab = "Total N",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -0, label.x = )

```

## Linear multiple regression models for all measures of N
# take away message: no models are significant. In otherwords, there is no evidence that testing N at a 
# particular MAT has any relationship to yeild.
```{r, echo = TRUE}
as.factor(tomatofun_yield_N_NONA$"Total.Yield.per.sq.ft")
as.factor(tomatofun_yield_N_NONA$"Sample.Period")
as.factor(tomatofun_yield_N_NONA$"Nitrate.N..ppm..SME")
as.factor(tomatofun_yield_N_NONA$"Nitrate.N..ppm..MM")
as.factor(tomatofun_yield_N_NONA$"TN.....Leaf.Analysis")
```

```{r}
#run model
M1 <- lm(tomatofun_yield_N_NONA$"Total.Yield.per.sq.ft" ~ tomatofun_yield_N_NONA$"Nitrate.N..ppm..SME" + tomatofun_yield_N_NONA$"Sample.Period", data = tomatofun_yield_N_NONA, method = "qr")

#check assumptions
mean(M1$residuals)        #Mean is <1, which is good.
plot(M1)                  #Residuals vs fitted indicate homoscedasticity. Will need to rerun model. Points are normally distributed.
acf(M1$residuals)         #No autocorrelation of residuals.
lmtest::dwtest(M1)        #No autocorrelation of residuals.
cor.test(tomatofun_yield_N_NONA$Nitrate.N..ppm..SME, M1$residuals)   #X variables and residuals are not correlated.
var(tomatofun_yield_N_NONA$Nitrate.N..ppm..SME)   #Variability is positive.
vif(M1)                   #Low GIF (<4) means low chance of multi-collinearity. 
gvlma::gvlma(M1)          #Shortcut for checking assumptions. 
summary(M1)

# if the model had been successful, I would take the time to remove the outliers and re-run, but since the model is not 
# significant, I think it would be a waste of time.

M2 <- lm(tomatofun_yield_N_NONA$"Total.Yield.per.sq.ft" ~ tomatofun_yield_N_NONA$"Nitrate.N..ppm..MM" + tomatofun_yield_N_NONA$"Sample.Period", data = tomatofun_yield_N_NONA, method = "qr")
gvlma::gvlma(M2)
summary(M2)

M3 <- lm(tomatofun_yield_N_NONA$"Total.Yield.per.sq.ft" ~ tomatofun_yield_N_NONA$TN.....Leaf.Analysis + tomatofun_yield_N_NONA$"Sample.Period", data = tomatofun_yield_N_NONA, method = "qr")
gvlma::gvlma(M3)
summary(M3)

```
head(data)
# P
```{r}
tomatofun_yield_P <- subset(data, select = c("Total.Yield.per.sq.ft", "Sample.Period", "Phosphorus..ppm..SME", "Phosphorus..lb.Ac..MM", "P.....Leaf.Analysis"))

tomatofun_yield_P_NONA <- na.omit(tomatofun_yield_P)

#set comparisons for error plots
my_comparisons <- list(c("Pre-Plant", "1-MAT"), c("Pre-Plant", "2-MAT"), c("Pre-Plant", "3-MAT"), c("Pre-Plant", "4-MAT"), c("Pre-Plant", "5-MAT"))
my_comparisonsb <- list(c("1-MAT", "2-MAT"), c("1-MAT", "3-MAT"), c("1-MAT", "4-MAT"), c("1-MAT", "5-MAT"))

#correct order of sample periods
tomatofun_yield_P_NONA$"Sample.Period" <- factor(tomatofun_yield_P_NONA$"Sample.Period", levels = c('Pre-Plant', 
                                                                                           '1-MAT', 
                                                                                            '2-MAT', 
                                                                                            '3-MAT', 
                                                                                            '4-MAT', 
                                                                                            '5-MAT'))

ggerrorplot(tomatofun_yield_P_NONA, x = "Sample.Period", 
            y = "Phosphorus..ppm..SME",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and SME phosphorus",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-5, 19),
            legend = "bottom",
            xlab = "sample period",
            ylab = "SME Phospohorus",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -5, label.x = )

ggerrorplot(tomatofun_yield_P_NONA, x = "Sample.Period", 
            y = "Phosphorus..lb.Ac..MM",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and MM phosphorus",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-1000, 2000),
            legend = "bottom",
            xlab = "sample period",
            ylab = "MM Phosphorus",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -1000, label.x = )

ggerrorplot(tomatofun_yield_P_NONA, x = "Sample.Period", 
            y = "P.....Leaf.Analysis",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and total P(leaf)",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-.25, 1),
            legend = "bottom",
            xlab = "sample period",
            ylab = "Total P",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -.2, label.x = )
```
## Linear multiple regression models for all measures of P
# take away message: no models are significant. In otherwords, there is no evidence that testing P at a 
# particular MAT has any relationship to yeild.

```{r}

M4 <- lm(tomatofun_yield_P_NONA$"Total.Yield.per.sq.ft" ~ tomatofun_yield_P_NONA$"Phosphorus..ppm..SME" + tomatofun_yield_P_NONA$"Sample.Period", data = tomatofun_yield_P_NONA, method = "qr")
gvlma::gvlma(M4)
summary(M2)

M5 <- lm(tomatofun_yield_P_NONA$"Total.Yield.per.sq.ft" ~ tomatofun_yield_P_NONA$"Phosphorus..lb.Ac..MM" + tomatofun_yield_P_NONA$"Sample.Period", data = tomatofun_yield_P_NONA, method = "qr")
gvlma::gvlma(M5)
summary(M5) #model is significant, however the only coefficient that is sig is the Ph.MM measure (not any sampling period)


```