---
title: "Tomato nutrient study correlations"
author: "Rachel Schattman and Becky Maden"
date: "November 7, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library
```{r}
library(dplyr)
```

## Start script


```{r}
script_path <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018"
in_dir <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018"
out_dir <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/output"
data <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/TomatoHT_Data_CHARTS_10.31.18_nutrientlevels.csv"))
```

## Create dataframe w/variables of interest, omit rows with NA

```{r}

tomatofun_nitrate <- subset(data, select = c("Nitrate.N..ppm..SME", "Nitrate.N..ppm..MM", "TN.....Leaf.Analysis"))
tomatofun_nitrateNONA <- na.omit(tomatofun_nitrate) 

```
## Check distribution - all are non normal
```{r}
shapiro.test(tomatofun_nitrateNONA$Nitrate.N..ppm..SME)
shapiro.test(tomatofun_nitrateNONA$Nitrate.N..ppm..MM)
shapiro.test(tomatofun_nitrateNONA$TN.....Leaf.Analysis)
```
## Correlation tests

```{r}

cor(tomatofun_nitrateNONA, method = "spearman")

cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..SME, y = tomatofun_nitrateNONA$Nitrate.N..ppm..MM, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..SME, y = tomatofun_nitrateNONA$TN.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..MM, y = tomatofun_nitrateNONA$TN.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)

```
## plot SME and MM correlations
```{r}

plot(tomatofun_nitrateNONA$Nitrate.N..ppm..SME, tomatofun_nitrateNONA$Nitrate.N..ppm..MM, type = "p", main = "Nitrate SME and MM, p < 0.001", xlab = "SME PPM", ylab = "MM PPM")
abline(lm(tomatofun_nitrateNONA$Nitrate.N..ppm..SME~tomatofun_nitrateNONA$Nitrate.N..ppm..MM), col = "red") # regression line
```
## plot MM and TN correlations
```{r}
plot(tomatofun_nitrateNONA$Nitrate.N..ppm..MM, tomatofun_nitrateNONA$TN.....Leaf.Analysis, type = "p", main = "Nitrate MM and TN Leaf, p < 0.05", xlab = "MM PPM", ylab = "TN leaf")
abline(lm(tomatofun_nitrateNONA$Nitrate.N..ppm..MM~tomatofun_nitrateNONA$TN.....Leaf.Analysis), col = "red") # regression line
