---
title: "Tomato nutrient study correlations, 2nd version"
author: "Rachel Schattman"
date: "November 27, 2018"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library
```{r}
library(dplyr)
library(ggpubr)
library(ggplot2)
library(car)
library(gvlma)
library(reshape2)
```
## Questions of interest

1.	The effects of cumulative nutrient applications on leaf tissue and soil nutrient levels the month after. (MASS only)
2.	The effects of soil test in any given month and the leaf tissue samples of the following month.
3.	The effects of diseases on yield (H1: powdery mildew will have the biggest effect)
4.	The effect of compaction on yield
5.	The effect of plant density on yield
6.	The effect of planting date on yield


## Start script
# Data preparation

```{r}
script_path <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV"
in_dir <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV"
out_dir <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/output/CSV"
```

# read in datasets
```{r}
FARMS <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/Farms.csv"))
NUTADDED <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/NutAdd.csv"))
NUTLEVELS <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/NutLevel.csv"))
PESTS <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/Pests.csv"))
SPACEYIELD <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/CSV/SpaceYield.csv"))
```

## Merge dataframes
```{r}
NUTS <-merge(NUTADDED, SPACEYIELD)
NUTS2 <- merge(NUTS, NUTLEVELS)
```


# Question 2: The effects of soil test in any given month and the leaf tissue samples of the following month.

## Test variables for normal distribution
### Note to self: figure out this whole batch processing business. 
```{r}
#shapiro.test(NUTLEVELS$TMY)   #not normal
#shapiro.test(NUTLEVELS$MYsqFT)#not normal
#shapiro.test(NUTLEVELS$NSME)  #not normal
#shapiro.test(NUTLEVELS$ASME)  #not normal
#shapiro.test(NUTLEVELS$PSME)  #not normal
#shapiro.test(NUTLEVELS$KSME)  #not normal
#shapiro.test(NUTLEVELS$MagSME)  #NORMAL DIST
#shapiro.test(NUTLEVELS$CaSME) #not normal
#shapiro.test(NUTLEVELS$AlSME) #not normal
#shapiro.test(NUTLEVELS$BoSME) #not normal
#shapiro.test(NUTLEVELS$CoSME) #not normal
#shapiro.test(NUTLEVELS$FeSME)   #NORMAL DIST

#NUTLEVELS$MangSME <- as.numeric(NUTLEVELS$MangSME) #transform into numberic
#shapiro.test(NUTLEVELS$MangSME) #NORMAL DIST

#shapiro.test(NUTLEVELS$SoSME) #not normal
#shapiro.test(NUTLEVELS$SuSME) #not normal
#shapiro.test(NUTLEVELS$ZSME)  #not normal
#shapiro.test(NUTLEVELS$NMM)   #not normal
#shapiro.test(NUTLEVELS$PMM)   #not normal
#shapiro.test(NUTLEVELS$KMM)     #NORMAL DIST
#shapiro.test(NUTLEVELS$CaMM)    #NORMAL DIST
#shapiro.test(NUTLEVELS$MaMM)  #not normal
#shapiro.test(NUTLEVELS$SuMM)  #not normal
#shapiro.test(NUTLEVELS$BoMM)  #not normal
#shapiro.test(NUTLEVELS$CoMM)  #not normal
#shapiro.test(NUTLEVELS$FeMM)    #NORMAL DIST
#shapiro.test(NUTLEVELS$MangMM)#not normal
#shapiro.test(NUTLEVELS$ZMM)   #not normal

#NUTLEVELS$TNLA <- as.numeric(NUTLEVELS$TNLA)
#shapiro.test(NUTLEVELS$TNLA)  #not normal

#NUTLEVELS$CALA <- as.numeric(NUTLEVELS$CALA)
#shapiro.test(NUTLEVELS$CALA)  #not normal

#NUTLEVELS$KLA <- as.numeric(NUTLEVELS$KLA)
#shapiro.test(NUTLEVELS$KLA)   #not normal

#NUTLEVELS$MgLA <- as.numeric(NUTLEVELS$MgLA)
#shapiro.test(NUTLEVELS$MgLA)  #not normal

#NUTLEVELS$PLA <- as.numeric(NUTLEVELS$PLA)
#shapiro.test(NUTLEVELS$PLA)   #not normal

#NUTLEVELS$AlLA <- as.numeric(NUTLEVELS$AlLA)
#shapiro.test(NUTLEVELS$AlLA)  #not normal

#NUTLEVELS$BLA <- as.numeric(NUTLEVELS$BLA)
#shapiro.test(NUTLEVELS$BLA)   #not normal

#NUTLEVELS$CuLA <- as.numeric(NUTLEVELS$CuLA)
#shapiro.test(NUTLEVELS$CuLA)  #not normal

#NUTLEVELS$FeLA <- as.numeric(NUTLEVELS$FeLA)
#shapiro.test(NUTLEVELS$FeLA)  #not normal

#NUTLEVELS$MnLA <- as.numeric(NUTLEVELS$MnLA)
#shapiro.test(NUTLEVELS$MnLA)  #not normal

#NUTLEVELS$ZLA <- as.numeric(NUTLEVELS$ZLA)
#shapiro.test(NUTLEVELS$ZLA)   #not normal
```

## Subset modified morgans (MM), saturdated media (SM) and leaf tissue (LT) tests by month after transplan (MX in SampPer). MM and SM are for same month, leaf tissue is for following month.
### First month soil, second month leaf
```{r}
# subset 1M rows, MM and SM columns
MM_SM_M1 <- NUTLEVELS[which(NUTLEVELS$SampPer == "1M"), names(NUTLEVELS) %in% c("ID", "TMY", "MYsqFT", "NSME", "ASME", "PSME", "KSME", "MagSME", "CaSME", "AlSME", "BoSME", "CoSME", "FeSME", "MangSME", "SoSME", "SuSME", "ZSME", "NMM", "PMM", "KMM", "CaMM", "MaMM",  "SuMM", "BoMM", "CoMM", "FeMM", "MangMM", "ZMM")] 

# subset 2M rows, LT columns
LA_M2 <- NUTLEVELS[which(NUTLEVELS$SampPer == "2M"), names(NUTLEVELS) %in% c("ID", "TNLA", "CALA", "KLA", "MgLA", "PLA", "AlLA", "BLA", "CuLA", "FeLA", "MnLA",  "ZLA")]

# merge MM_SM_M1 and LA_M2
M1_M2 <- merge(MM_SM_M1, LA_M2)

# some columns need to be tranformed into numberic vectors 
# M1_M2 <- sapply(M1_M2, as.numeric)
# is.numeric(M1_M2)
```
### Second month soil, third month leaf
```{r}
# subset 2M rows, MM and SM columns
MM_SM_M2 <- NUTLEVELS[which(NUTLEVELS$SampPer == "2M"), names(NUTLEVELS) %in% c("ID", "TMY", "MYsqFT", "NSME", "ASME", "PSME", "KSME", "MagSME", "CaSME", "AlSME", "BoSME", "CoSME", "FeSME", "MangSME", "SoSME", "SuSME", "ZSME", "NMM", "PMM", "KMM", "CaMM", "MaMM",  "SuMM", "BoMM", "CoMM", "FeMM", "MangMM", "ZMM")] 

# subset 3M rows, LT columns
LA_M3 <- NUTLEVELS[which(NUTLEVELS$SampPer == "3M"), names(NUTLEVELS) %in% c("ID", "TNLA", "CALA", "KLA", "MgLA", "PLA", "AlLA", "BLA", "CuLA", "FeLA", "MnLA",  "ZLA")]

# merge MM_SM_M2 and LA_M3
M2_M3 <- merge(MM_SM_M2, LA_M3)
```
### Third month soil, fourth month leaf
```{r}
# subset 3M rows, MM and SM columns
MM_SM_M3 <- NUTLEVELS[which(NUTLEVELS$SampPer == "3M"), names(NUTLEVELS) %in% c("ID", "TMY", "MYsqFT", "NSME", "ASME", "PSME", "KSME", "MagSME", "CaSME", "AlSME", "BoSME", "CoSME", "FeSME", "MangSME", "SoSME", "SuSME", "ZSME", "NMM", "PMM", "KMM", "CaMM", "MaMM",  "SuMM", "BoMM", "CoMM", "FeMM", "MangMM", "ZMM")] 

# subset 4M rows, LT columns
LA_M4 <- NUTLEVELS[which(NUTLEVELS$SampPer == "4M"), names(NUTLEVELS) %in% c("ID", "TNLA", "CALA", "KLA", "MgLA", "PLA", "AlLA", "BLA", "CuLA", "FeLA", "MnLA",  "ZLA")]

# merge MM_SM_M3 and LA_M4
M3_M4 <- merge(MM_SM_M3, LA_M4)
```
### Fourth month soil, fifth month leaf
```{r}
# subset 3M rows, MM and SM columns
MM_SM_M4 <- NUTLEVELS[which(NUTLEVELS$SampPer == "4M"), names(NUTLEVELS) %in% c("ID", "TMY", "MYsqFT", "NSME", "ASME", "PSME", "KSME", "MagSME", "CaSME", "AlSME", "BoSME", "CoSME", "FeSME", "MangSME", "SoSME", "SuSME", "ZSME", "NMM", "PMM", "KMM", "CaMM", "MaMM",  "SuMM", "BoMM", "CoMM", "FeMM", "MangMM", "ZMM")] 

# subset 4M rows, LT columns
LA_M5 <- NUTLEVELS[which(NUTLEVELS$SampPer == "5M"), names(NUTLEVELS) %in% c("ID", "TNLA", "CALA", "KLA", "MgLA", "PLA", "AlLA", "BLA", "CuLA", "FeLA", "MnLA",  "ZLA")]

# merge MM_SM_M4 and LA_M5
M4_M5 <- merge(MM_SM_M3, LA_M4)
```

## Regress leaf tissue analysis on SME and MM for N
```{r}
# N
# 1MAT soil, 2MAT leaf --> No sig correlation

cor.test(x = M1_M2$NSME, y = M1_M2$TNLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M1_M2$TNLA), exact = NULL)
cor.test(x = M1_M2$NMM, y = M1_M2$TNLA, alternative = "two.sided", method = "spearman", exact = NULL)

# 2MAT soil, 3MAT leaf --> No sig correlation

cor.test(x = M2_M3$NSME, y = M2_M3$TNLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M2_M3$TNLA), exact = NULL)
cor.test(x = M2_M3$NMM, y = M2_M3$TNLA, alternative = "two.sided", method = "spearman", exact = NULL)

# 3MAT soil, 4MAT leaf --> SME not sig correlated. MM is significantly correlated at the 0.05 level. However, linear model is not significant.

cor.test(x = M3_M4$NSME, y = M3_M4$TNLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M3_M4$TNLA), exact = NULL)
cor.test(x = M3_M4$NMM, y = M3_M4$TNLA, alternative = "two.sided", method = "spearman", exact = NULL)

mod1 <- lm(NSME~TNLA, data=M3_M4)

summary(mod1)

plot(M3_M4$NMM, M3_M4$TNLA, type = "p", main = "TN leaf (4 MAT) ~ N Saturated Media (3 MAT)", xlab = "Modified Morgans ppm", ylab = "Total N leaf analysis ppm")
abline(lm(M3_M4$TNLA~M3_M4$NMM), col = "red") # regression line
 
# 4MAT soil, 5MAT leaf --> SME not sig correlated. MM is significantly correlated at the 0.05 level. However, linear model is not significant.

cor.test(x = M4_M5$NSME, y = M4_M5$TNLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M4_M5$TNLA), exact = NULL)
cor.test(x = M4_M5$NMM, y = M4_M5$TNLA, alternative = "two.sided", method = "spearman", exact = NULL)

mod2 <- lm(NSME~TNLA, data=M4_M5)

summary(mod2)

plot(M4_M5$NMM, M4_M5$TNLA, type = "p", main = "TN leaf (5 MAT) ~ N Saturated Media (4 MAT)", xlab = "Modified Morgans ppm", ylab = "Total N leaf analysis ppm")
```

## Regress leaf tissue analysis on SME and MM for P
```{r}
# P
# 1MAT soil, 2MAT leaf --> No sig correlation

cor.test(x = M1_M2$PSME, y = M1_M2$PLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M1_M2$PLA), exact = NULL)
cor.test(x = M1_M2$PMM, y = M1_M2$PLA, alternative = "two.sided", method = "spearman", exact = NULL)

# 2MAT soil, 3MAT leaf --> No sig correlation

cor.test(x = M2_M3$PSME, y = M2_M3$PLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M2_M3$PLA), exact = NULL)
cor.test(x = M2_M3$PMM, y = M2_M3$PLA, alternative = "two.sided", method = "spearman", exact = NULL)

# 3MAT soil, 4MAT leaf --> No sig relationship

cor.test(x = M3_M4$PSME, y = M3_M4$PLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M3_M4$PLA), exact = NULL)
cor.test(x = M3_M4$PMM, y = M3_M4$PLA, alternative = "two.sided", method = "spearman", exact = NULL)
 
# 4MAT soil, 5MAT leaf --> No sig relationship

cor.test(x = M4_M5$PSME, y = M4_M5$PLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M4_M5$PLA), exact = NULL)
cor.test(x = M4_M5$PMM, y = M4_M5$PLA, alternative = "two.sided", method = "spearman", exact = NULL)
```
```{r}
# K
# 1MAT soil, 2MAT leaf --> No sig correlation

cor.test(x = M1_M2$KSME, y = M1_M2$KLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M1_M2$KLA), exact = NULL)
cor.test(x = M1_M2$KMM, y = M1_M2$KLA, alternative = "two.sided", method = "spearman", exact = NULL)

# 2MAT soil, 3MAT leaf --> No sig correlation

cor.test(x = M2_M3$KSME, y = M2_M3$KLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M2_M3$KLA), exact = NULL)
cor.test(x = M2_M3$KMM, y = M2_M3$KLA, alternative = "two.sided", method = "spearman", exact = NULL)

# 3MAT soil, 4MAT leaf --> No sig relationship

cor.test(x = M3_M4$KSME, y = M3_M4$KLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M3_M4$KLA), exact = NULL)
cor.test(x = M3_M4$KMM, y = M3_M4$KLA, alternative = "two.sided", method = "spearman", exact = NULL)
 
# 4MAT soil, 5MAT leaf --> No sig relationship

cor.test(x = M4_M5$KSME, y = M4_M5$KLA, alternative = "two.sided", method = "spearman", na.action = na.omit(M4_M5$KLA), exact = NULL)
cor.test(x = M4_M5$KMM, y = M4_M5$KLA, alternative = "two.sided", method = "spearman", exact = NULL)
```


