---
title: "Tomato nutrient study correlations"
author: "Rachel Schattman and Becky Maden"
date: "November 7, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load library
```{r}
library(dplyr)
library(ggpubr)
```

## Start script


```{r}
script_path <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018"
in_dir <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018"
out_dir <- "C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/output"
data <- data.frame(read.csv("C:/Users/rschattman/Documents/Research/Greenhouse_tomatoes_2018/Nutrient_levels_yield_NH_VT_MA_11_14_18.csv"))
```

# Nitrate
## Create dataframe w/variables of interest for Nitrate only, omit rows with NA


```{r}

tomatofun_nitrate <- subset(data, select = c("Nitrate.N..ppm..SME", "Nitrate.N..ppm..MM", "TN.....Leaf.Analysis"))
tomatofun_nitrateNONA <- na.omit(tomatofun_nitrate) 

```

## Check distribution - all are non normal

```{r}
shapiro.test(tomatofun_nitrateNONA$Nitrate.N..ppm..SME)
shapiro.test(tomatofun_nitrateNONA$Nitrate.N..ppm..MM)
shapiro.test(tomatofun_nitrateNONA$TN.....Leaf.Analysis)
```
## Correlation tests

```{r}

cor(tomatofun_nitrateNONA, method = "spearman")

cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..SME, y = tomatofun_nitrateNONA$Nitrate.N..ppm..MM, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..SME, y = tomatofun_nitrateNONA$TN.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_nitrateNONA$Nitrate.N..ppm..MM, y = tomatofun_nitrateNONA$TN.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)

```
## plot SME and MM correlations
```{r}
?plot
plot(tomatofun_nitrateNONA$Nitrate.N..ppm..SME, tomatofun_nitrateNONA$Nitrate.N..ppm..MM, type = "p", main = "Nitrate SME and MM, p < 0.001", xlab = "SME PPM", ylab = "MM PPM")
abline(lm(tomatofun_nitrateNONA$Nitrate.N..ppm..SME~tomatofun_nitrateNONA$Nitrate.N..ppm..MM), col = "red") # regression line
```
# Phosphorus
## Create dataframe w/variables of interest for P only, omit rows with NA

```{r}
tomatofun_P <- subset(data, select = c("Phosphorus..ppm..SME", "Phosphorus..lb.Ac..MM", "P.....Leaf.Analysis"))
tomatofun_P_NONA <- na.omit(tomatofun_P) 

```

## Check distribution 

```{r}
shapiro.test(tomatofun_P_NONA$Phosphorus..ppm..SME)
shapiro.test(tomatofun_P_NONA$Phosphorus..lb.Ac..MM)
shapiro.test(tomatofun_P_NONA$P.....Leaf.Analysis)
```

## Correlation tests

```{r}

cor(tomatofun_P_NONA, method = "spearman")

cor.test(x = tomatofun_P_NONA$Phosphorus..ppm..SME, y = tomatofun_P_NONA$Phosphorus..lb.Ac..MM, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_P_NONA$Phosphorus..ppm..SME, y = tomatofun_P_NONA$P.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_P_NONA$Phosphorus..lb.Ac..MM, y = tomatofun_P_NONA$P.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)

```

## plot SME and MM correlations
```{r}

plot(tomatofun_P_NONA$Phosphorus..lb.Ac..MM, tomatofun_P_NONA$Phosphorus..ppm..SME, type = "p", main = "Phosphorus SME and MM, p < 0.001", xlab = "MM PPM", ylab = "SME PPM")

abline(lm(tomatofun_P_NONA$Phosphorus..ppm..SME~tomatofun_P_NONA$Phosphorus..lb.Ac..MM), col = "red") # regression line
```

# Potassium
## Create dataframe w/variables of interest for P only, omit rows with NA
head(data)

```{r}
tomatofun_K <- subset(data, select = c("Potassium..ppm..SME", "Potassium..lb.Ac..MM", "K.....Leaf.Analysis"))
tomatofun_K_NONA <- na.omit(tomatofun_K) 

```

## Check distribution 

```{r}
shapiro.test(tomatofun_K_NONA$Potassium..ppm..SME)
shapiro.test(tomatofun_K_NONA$Potassium..lb.Ac..MM)
shapiro.test(tomatofun_K_NONA$K.....Leaf.Analysis)
```

```{r}

cor(tomatofun_K_NONA, method = "spearman")

cor.test(x = tomatofun_K_NONA$Potassium..ppm..SME, y = tomatofun_K_NONA$Potassium..lb.Ac..MM, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_K_NONA$Potassium..ppm..SME, y = tomatofun_K_NONA$K.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)
cor.test(x = tomatofun_K_NONA$Potassium..lb.Ac..MM, y = tomatofun_K_NONA$K.....Leaf.Analysis, alternative = "two.sided", method = "spearman", exact = NULL)

```

## plot SME and MM correlations
```{r}

plot(tomatofun_K_NONA$Potassium..ppm..SME, tomatofun_K_NONA$Potassium..lb.Ac..MM, type = "p", main = "Potassium SME and MM, p < 0.001", xlab = "SME PPM", ylab = "MM PPM")
abline(lm(tomatofun_K_NONA$Potassium..lb.Ac..MM~tomatofun_K_NONA$Potassium..ppm..SME), col = "red") # regression line
```

# P SME and pH MM
## Create dataframe w/variables of interest, omit rows with NA
head(data)
```{r}
tomatofun_P_pH <- subset(data, select = c("Phosphorus..ppm..SME", "Soil.pH"))
tomatofun_P_pH_NONA <- na.omit(tomatofun_P_pH) 
```

```{r}

cor(tomatofun_P_pH_NONA, method = "spearman")

cor.test(x = tomatofun_P_pH_NONA$Phosphorus..ppm..SME, y = tomatofun_P_pH_NONA$Soil.pH, alternative = "two.sided", method = "spearman", exact = NULL)

plot(tomatofun_P_pH_NONA$Phosphorus..ppm..SME, tomatofun_P_pH_NONA$Soil.pH, type = "p", main = "Phosphorus SME and pH, p < 0.001", xlab = "SME PPM", ylab = "pH")
abline(lm(tomatofun_P_pH_NONA$Soil.pH~tomatofun_P_pH_NONA$Phosphorus..ppm..SME), col = "red") # regression line
```

Take home message: No significant difference between Nitrate levels (any test) by sample period. 
```{r}

tomatofun_yield_N <- subset(data, select = c("Total.Yield.per.sq.ft", "Sample.Period", "Nitrate.N..ppm..SME", "Nitrate.N..ppm..MM", "TN.....Leaf.Analysis"))

tomatofun_yield_N_NONA <- na.omit(tomatofun_yield_N)

#set comparisons for error plots
my_comparisons <- list(c("Pre-Plant", "1-MAT"), c("Pre-Plant", "2-MAT"), c("Pre-Plant", "3-MAT"), c("Pre-Plant", "4-MAT"), c("Pre-Plant", "5-MAT"))
my_comparisonsb <- list(c("1-MAT", "2-MAT"), c("1-MAT", "3-MAT"), c("1-MAT", "4-MAT"), c("1-MAT", "5-MAT"))

#correct order of sample periods
tomatofun_yield_N_NONA$"Sample.Period" <- factor(tomatofun_yield_N_NONA$"Sample.Period", levels = c('Pre-Plant', 
                                                                                           '1-MAT', 
                                                                                            '2-MAT', 
                                                                                            '3-MAT', 
                                                                                            '4-MAT', 
                                                                                            '5-MAT'))

ggerrorplot(tomatofun_yield_N_NONA, x = "Sample.Period", 
            y = "Nitrate.N..ppm..SME",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and SME nitrate",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-500, 1000),
            legend = "bottom",
            xlab = "sample period",
            ylab = "SME Nitrate",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -250, label.x = )

ggerrorplot(tomatofun_yield_N_NONA, x = "Sample.Period", 
            y = "Nitrate.N..ppm..MM",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and MM nitrate",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-500, 1000),
            legend = "bottom",
            xlab = "sample period",
            ylab = "MM Nitrate",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -250, label.x = )

ggerrorplot(tomatofun_yield_N_NONA, x = "Sample.Period", 
            y = "TN.....Leaf.Analysis",
            combine = FALSE, merge = TRUE,
            desc_stat = "mean_ci",  
            color = "black",
            palette = "npg",
            title = "Sampling period and total N(leaf)",
            add = "violin", add.params = list(color = "darkgray", fill="lightgray"),
            ylim = c(-2, 8),
            legend = "bottom",
            xlab = "sample period",
            ylab = "Total N",
            orientation = "vertical") + 
            stat_compare_means(comparisons = my_comparisonsb) +
            stat_compare_means(label.y = -0, label.x = )

```